Backscatter Modulation Design for Symbiotic Radio Networks

### 2022-01-26 main.tex
:pencil: Update thresholding proof

			\begin{theorem}
				If any $L$-input continuous-output channel follows Erlang distribution~\eqref{eq:energy_distribution}, then the optimal number of thresholds is $L+1$ and the capacity-achieving decision region $j$ has convex structure $\mathcal{R}_j^\star = [t_{j-1}^\star, t_j^\star)$.
			\end{theorem}

			\begin{proof}
				For any finite number of thresholds $S \ge L+1$, let $\boldsymbol{t} \triangleq [t_0,\ldots,t_S]^T \in \mathbb{R}_{+}^{S \times 1}$ be the thresholding vector where $t_{s-1} < t_s$, $\forall s \in \mathcal{S} \triangleq \{1,\ldots,S\}$.\footnote{We consider strict inequality in the definition, as otherwise $S$ and $\boldsymbol{t}$ could be non-unique for a given thresholding scheme.} Denote the decision partition $s$ as $R_s \triangleq [t_{s-1},t_s)$. Without loss of generality, we assume
				% $L \mid S$ and
				$\mathcal{R}_j = \bigcup_{s \equiv j \pmod L} R_s$.\footnote{The proof holds for general mapping $R_s \mapsto \mathcal{R}_j$, and we focus on this specific case for the ease of notation.} The thresholding design problem can be expressed as
				\begin{maxi!}
					{\scriptstyle{\boldsymbol{t}}}{I_{\mathrm{B}}(c_{\mathcal{K}};z)}{\label{op:thresholding}}{\label{ob:backscatter_mutual_information}}
					\addConstraint{t_{s-1}}{< t_s,}{\quad \forall s \in \mathcal{S}.}{\label{co:strict_inequality}}
				\end{maxi!}

				Problem~\eqref{op:thresholding} is intricate due to the strict inequality constraint~\eqref{co:strict_inequality}. Following \cite{Nguyen2020}, we first relax it to the convex counterpart, then discard the solutions that violate the original constraints. The Lagrangian function for the relaxed problem is
				\begin{equation}
					L(\boldsymbol{t},\boldsymbol{\mu}) = - I_{\mathrm{B}}(c_{\mathcal{K}};z) + \sum_{s \in \mathcal{S}} \mu_s (t_{s-1} - t_s),
					% = \sum_{j \in \mathcal{L}} P(\bar{z}_j) \log P(\bar{z}_j) - \sum_{i \in \mathcal{L}} \sum_{j \in \mathcal{L}}
				\end{equation}
				where $\mu_s$ is the Lagrange multiplier associated with the non-strict version of \eqref{co:strict_inequality} and $\boldsymbol{\mu} \triangleq [\mu_1,\ldots,\mu_S]^T \in \mathbb{R}^{S \times 1}$. The corresponding \gls{kkt} conditions on the optimal primal and dual solutions are, $\forall s \in \mathcal{S}$,
				\begin{subequations}
					\label{eq:kkt_thresholding}
					\begin{equation}
						- \nabla_{t_s^\star} I^\star_{\mathrm{B}}(c_{\mathcal{K}};z) + \mu_{s-1}^\star - \mu_s^\star = 0,
						\label{eq:stationarity}
					\end{equation}
					\begin{equation}
						\mu_s^\star \ge 0,
						\label{eq:dual_feasibility}
					\end{equation}
					\begin{equation}
						\mu_s^\star (t_{s-1}^\star - t_s^\star) = 0.
						\label{eq:complementary_slackness}
					\end{equation}

				\end{subequations}
				Due to the strict inequality constraint~\eqref{co:strict_inequality}, conditions \eqref{eq:dual_feasibility} and \eqref{eq:complementary_slackness} together imply $\mu_s^\star = 0$, $\forall s \in \mathcal{S}$. For \gls{dmtc} based on energy detection~\eqref{eq:energy_distribution}, it is trivial to conclude $t_0^\star = 0$ and $t_S^\star = \infty$. Hence, we focus on $s' \in \mathcal{S} \setminus \{S\}$ and \eqref{eq:kkt_thresholding} reduces to
				\begin{equation}
					\nabla_{t_{s'}^\star} I^\star_{\mathrm{B}}(c_{\mathcal{K}};z) = 0,
				\end{equation}
				that is, for $j \in \mathcal{L}$ and $j \equiv s' \pmod L$,
				\begin{equation}
					\sum_l a_{l,j} f(t_{s'}^\star \mid x_l) = 0,
					\label{eq:kkt_thresholding_explicit}
				\end{equation}
				where
				\begin{equation}
					a_{l,j} =
					\begin{cases}
						P(x_l) \log \frac{P(x_l \mid z_{j+1}^\star)}{P(x_l \mid z_j^\star)}, & j \in \mathcal{L} \setminus \{L\},\\
						P(x_l) \log \frac{P(x_l \mid z_{1}^\star)}{P(x_l \mid z_j^\star)}, & j = L.
					\end{cases}
				\end{equation}

				Since the optimal backward channel quantizer is convex and separates each pair of posterior distribution by a hyperplane~\cite{Nguyen2020a}, the sequence $(a_{l,j})_{l=1}^{L}$ only change sign once for a given $j$. Besides, we notice the LHS of~\eqref{eq:kkt_thresholding_explicit} is a generalized Dirichlet polynomial, as $f(t_{s'}^\star \mid x_l)$ follows Erlang distribution~\eqref{eq:energy_distribution}. By Descartes' rule of signs~\cite{Jameson2006}, for each $j \in \mathcal{L}$, $t_s^\star$ has at most 1 real solution if $\sigma_l^2 \ne \sigma_{l'}^2$, $l,l' \in \mathcal{L}$.

>>>>

			\begin{theorem}
				For any $L$-input continuous-output channel with Erlang distribution~\eqref{eq:energy_distribution}, the number of thresholds required to achieve the \gls{dmtc} capacity is $L+1$ and the optimal decision region $j \in \mathcal{L} \triangleq \{1,\ldots,L\}$ has a convex structure $\mathcal{R}_j^\star = [t_{j-1}^\star, t_j^\star)$.
			\end{theorem}

			\begin{proof}
				Assume the optimal number of thresholds is $S+1$. Since $x \to z \to \bar{z}$ formulates a Markov chain, we need at least $L+1$ thresholds to construct $L$ valid detection regions and minimize the distortion between $x$ and $\bar{z}$. It suggests $S \ge L$. Let $\boldsymbol{t} \triangleq [t_0,\ldots,t_S]^T \in \mathbb{R}_{+}^{(S+1) \times 1}$ be the optimal thresholding vector where $t_{s-1} < t_s$, $\forall s \in \mathcal{S} \triangleq \{1,\ldots,S\}$.\footnote{We consider strict inequality in the definition, as otherwise $S$ and $\boldsymbol{t}$ could be non-unique for a given thresholding scheme.} Denote the decision partition $s$ as $R_s \triangleq [t_{s-1},t_s)$. Without loss of generality, we assume $\mathcal{R}_j = \bigcup_{s \equiv j \pmod L} R_s$.\footnote{The proof holds for general mapping $R_s \mapsto \mathcal{R}_j$, and we focus on this specific case for the ease of notation.} The thresholding design problem can be expressed as
				\begin{maxi!}
					{\scriptstyle{\boldsymbol{t}}}{I_{\mathrm{B}}(c_{\mathcal{K}};z)}{\label{op:thresholding}}{\label{ob:backscatter_mutual_information}}
					\addConstraint{t_{s-1}}{< t_s,}{\quad \forall s \in \mathcal{S}.}{\label{co:strict_inequality}}
				\end{maxi!}

				Problem~\eqref{op:thresholding} is intricate due to the strict inequality constraint~\eqref{co:strict_inequality}. Following \cite{Nguyen2020}, we first relax it to the convex counterpart, then discard the solutions that violate the original constraints. The Lagrangian function for the relaxed problem is
				\begin{equation}
					L(\boldsymbol{t},\boldsymbol{\mu}) = - I_{\mathrm{B}}(c_{\mathcal{K}};z) + \sum_{s \in \mathcal{S}} \mu_s (t_{s-1} - t_s),
				\end{equation}
				where $\mu_s$ is the Lagrange multiplier associated with the non-strict version of \eqref{co:strict_inequality} and $\boldsymbol{\mu} \triangleq [\mu_1,\ldots,\mu_S]^T \in \mathbb{R}^{S \times 1}$. The \gls{kkt} conditions on the optimal primal and dual solutions are, $\forall s \in \mathcal{S}$,
				\begin{subequations}
					\label{eq:kkt_thresholding}
					\begin{equation}
						- \nabla_{t_s^\star} I^\star_{\mathrm{B}}(c_{\mathcal{K}};z) + \mu_{s-1}^\star - \mu_s^\star = 0,
						\label{eq:stationarity}
					\end{equation}
					\begin{equation}
						\mu_s^\star \ge 0,
						\label{eq:dual_feasibility}
					\end{equation}
					\begin{equation}
						\mu_s^\star (t_{s-1}^\star - t_s^\star) = 0.
						\label{eq:complementary_slackness}
					\end{equation}

				\end{subequations}
				% For \gls{dmtc} based on energy detection~\eqref{eq:energy_distribution}, it is trivial to conclude $t_0^\star = 0$ and $t_S^\star = \infty$.
				Due to the strict inequality constraint~\eqref{co:strict_inequality}, conditions \eqref{eq:dual_feasibility} and \eqref{eq:complementary_slackness} together imply $\mu_s^\star = 0$, $\forall s \in \mathcal{S}$. Besides, it is trivial to conclude $t_0^\star = 0$ for our \gls{dmtc} based on energy detection. As such, the necessary optimality condition for $t_s^\star$, $s \in \mathcal{S}$ is
				\begin{equation}
					\nabla_{t_{s}^\star} I^\star_{\mathrm{B}}(c_{\mathcal{K}};z) = 0,
				\end{equation}
				namely
				\begin{equation}
					\sum_l a_{l,j} f(t_{s}^\star \mid x_l) = 0,
					\label{eq:kkt_thresholding_explicit}
				\end{equation}
				where $j \equiv s \pmod L$, $j \in \mathcal{L}$ and
				\begin{equation}
					a_{l,j} \triangleq
					\begin{cases}
						P(x_l) \log \frac{P(x_l \mid z_{j+1}^\star)}{P(x_l \mid z_j^\star)}, & j \in \mathcal{L} \setminus \{L\},\\
						P(x_l) \log \frac{P(x_l \mid z_{1}^\star)}{P(x_l \mid z_j^\star)}, & j = L.
					\end{cases}
				\end{equation}

				Since the optimal backward channel quantizer is convex and separates each pair of posterior distribution by a hyperplane~\cite{Nguyen2020a}, for a given $j$, the sequence $(a_{l,j})_{l=1}^{L}$ only change sign once. As $f(t_{s}^\star \mid x_l)$ follows Erlang distribution~\eqref{eq:energy_distribution}, the LHS of~\eqref{eq:kkt_thresholding_explicit} is a generalized Dirichlet polynomial and has at most one real solution by Descartes' rule of signs~\cite{Jameson2006}. In other words, each decision region $j \in \mathcal{L}$ requires at most one additional threshold, and therefore $S \le L$. This suggests $S = L$ completes the proof.

				% . By Descartes' rule of signs~\cite{Jameson2006}, for each $j \in \mathcal{L}$, $t_s^\star$ has at most 1 real solution if $\sigma_l^2 \ne \sigma_{l'}^2$, $\forall l,l' \in \mathcal{L}$.

				% In other words, based on $t_0^\star = 0$, each decision region $j \in \mathcal{L}$ requires at most one additional threshold, such that $S \le L+1$.
				% $s \equiv j \pmod L$
